---
title: "Data Set 4"
author: "Christen Parzych"
date: "9/11/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

As always, I first cleared the environment and loaded each of the required libraries: 

```{r}

rm(list=ls())

library(tidyverse)


library(haven)


library(readxl)

library(readr)
```

##1. Panel data for OECD countries. Save as file name 'oecd.Rdata.'


This is a delimited dataset, and I used the 'read.delim' function to load it. I named it to the environment as 'oecd.' It has 342 observations of 6 variables. I used the 'head' command to verify that it loaded cleanly, and saved it as 'oecd.Rdata.'

```{r}
oecd<-read.delim(file="http://www.wiley.com/legacy/wileychi/baltagi/supp/Gasoline.dat",header=TRUE,dec = " ")

read_delim(file="http://www.wiley.com/legacy/wileychi/baltagi/supp/Gasoline.dat",delim = " ")

head(oecd)

save(oecd,file="oecd.Rdata")

```

##2. U.S. National Election Survey data from 2004. Save as file name nes.RData.
http://www.uta.edu/faculty/story/DataSets.htm

We know that this data is in SPSS format because it is a '.sav' file. As a result, I will use the 'haven' command, or 'read_spss.' Once again we use the 'head' command to verify that all the data was read in correctly. There are 1,212 observations of 242 variables. 

Finally, I saved it as an R file. 

```{r}
nes<-read_spss("http://www.uta.edu/faculty/story/DataSets/NES2004.sav")

head(nes)

save(nes,file = "nes.Rdata")

```


##3. General social survey, student version. Save as file name gss.Rdata.
http://www.uta.edu/faculty/story/DataSets.htm

This is an Excel file, so I will use the 'read_Excel' function to bring the data in to the environment. I've named the dataset 'gss.' We need to use a true/false structure to tell R whether certain conditions are met. In this case, we want R to include the column names, so I can set the condition as 'true.'  

I then saved the file as requested. The dataset has 995 observations of 44 variables. 

```{r}
gss<-read_excel("GeneralSocialSurvey1996ExcelDataStudentVersion.xls",skip=0, col_names = TRUE)

head(gss)

save(gss,file = "gss.Rdata")
```

##4. Replication file for "STATISTICAL DISCRIMINATION OR PREJUDICE? A LARGE SAMPLE FIELD EXPERIMENT." Open up and save the mainData.csv file. Save it as maindata.Rdata. 

This is a comma-separated variable file, which is a data file delimited by commas.The file is found at the following address:

https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/26410&studyListingIndex=1_f1a615c2cbee97fb79c9d575ae73 

I downloaded the 'mainData.csv' file and saved it to my working directory (/~Box Sync/School/DataScience/myrepo1). I then used the 'read_csv' function to load the file into R, naming it 'main.'

I then verified that the file loaded cleanly by using the 'head' function and saved the file using 'write_csv.' This dataset contain 14,237 observations of 41 variables. The column names appear clear and descriptive, so I did not rename them. 

```{r}
main<-read_csv(file="mainData.csv")

head(main)

write_csv(main,path="maindata.Rdata")

```

##5. The Lalonde dataset, covering work experiences in the Panel Study of Income Dynamics (psid). Save as psid.RData http://users.nber.org/~rdehejia/data/psid_controls.txt 

This is a delimited dataset, and I used the 'read.delim' function to load it. I named it to the environment as 'psid.' It has 2,490 observations of 21 variables, though many  of the columns are NA, so I will need to clean that up. I will also add labels to each variable/column. 

Once I have eliminated the blank columns (and renamed the dataset 'psid2'), I am left with 2,490 observations of ten variables.

The instructions point us to the data labels on the website. They are as follows (left to right): treatment indicator (1 if treated, 0 if not treated), age, education, Black (1 if black, 0 otherwise), Hispanic (1 if Hispanic, 0 otherwise), married (1 if married, 0 otherwise), nodegree (1 if no degree, 0 otherwise), RE75 (earnings in 1975), and RE78 (earnings in 1978). The last variable is the outcome; other variables are pre-treatment.
I then saved it as 'psid.Rdata.'

```{r}
psid<-read.delim(file="http://users.nber.org/~rdehejia/data/psid_controls.txt",header=FALSE,sep = " ")

head(psid)

psid2<-psid[ ,-(c(1,2,4,6,8,10,12,14,16,18,20))]

names(psid2)<-c("treatment_indicator",
                "age",
                "education",
                "black",
                "hispanic",
                "married",
                "nodegree",
                "RE75",
                "RE78",
                "outcome")

head(psid2)

save(psid2,file = "psid.Rdata")
```